### 2011-08-11 nature_sci_rep/submission2/cover_letter.tex
ready to resubmit?

The philosophical notion of supervenience, colloquially, that there cannot be an A-difference without a corresponding B-difference, has received much attention in the philosophy literature; but relatively scant attention from the statistics and science communities.  We believe this is an unfortunate and unnecessary consequence of scientific reductionism.  In this work, we very briefly present the philosophical notion of supervenience canonized by Davidson (Experience and Theory, 1970).  We then rapidly launch into a statistical understanding